Arc Blanc: a real time ocean simulation framework

### 2025-02-03 paper.tex
Update appendix on "Evaluation of the Number of Iterations for Convergence"

\section{Evaluation of the Number of Iterations for Convergence}
\label{sec:evaluation_of_the_number_of_iterations_for_convergence}
The optimal number of iterations ($N_{\text{iter}}$) required for the iterative retrieval of the water height was determined through a series of tests. Wind speeds were varied from $0.1 \, \text{m/s}$ to $35.0 \, \text{m/s}$ in increments of $0.5 \, \text{m/s}$, and the convergence of water height was evaluated at $1000$ random positions over a $10^4 \, \text{m} \times 10^4 \, \text{m}$ domain.

The program iteratively computed the water height until the changes in height were smaller than a $1$ cm.

The results, as shown in Figure~\ref{fig:convergence_steps}, demonstrate that the required number of iterations increases with wind speed, stabilizing at approximately $N_{\text{iter}} = 4$ for higher wind speeds. This finding suggests that using $N_{\text{iter}} = 4$ ensures reliable convergence for all tested conditions while balancing computational efficiency.

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.9\textwidth]{Figures/results_cm.png}
    \caption{Mean number of steps required for convergence as a function of wind speed. Error bars represent the variance of iterations across $1000$ test points.}
    \label{fig:convergence_steps}
\end{figure}
%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%
\afterdoc
%%%%%%%%%%%%%%%
\end{document}

>>>>

\section{Evaluation of the Number of Iterations for Convergence}
\label{sec:evaluation_of_the_number_of_iterations_for_convergence}

The optimal number of iterations ($N_{\text{iter}}$) required for the iterative retrieval of water height was determined through two experiments, focusing on accuracy and computational efficiency.

The first experiment aimed to evaluate the number of iterations required for the water height computation to converge within a given precision. This test has been conducted on different wind speeds as it makes the height of water significantly. Wind speeds ranged from $0.1$ m/s to $35$ m/s in $0.5$ m/s increments, with convergence tested at $10^3$ random positions uniformly distributed over a $10^4\times 10^4$ m domain. The iterative process was repeated until the difference between consecutive height estimates was smaller than $0.01$ m. The results, shown in Figure~\ref{fig:convergence_steps}, indicate that the required number of iterations increases with wind speed but stabilizes at approximately $N_{\text{iter}} = 4$ for higher wind speeds.

The second experiment aimed to measure the computational cost associated with varying numbers of iterations. The execution time was assessed using a CUDA-based implementation, where the iterative retrieval of water height was performed for different values of $N_{\text{iter}}$. Each test was executed $10^3$ times on a GPU, and execution times were recorded across $10^5$ randomly sampled positions. The measurements included the minimum, maximum, mean, and variance of execution times for each iteration count. The results, presented in Figure~\ref{fig:execution_time}, show that while execution time increases with the number of iterations, the rate of increase diminishes after four iterations. This confirms that $N_{\text{iter}} = 4$ provides an optimal balance between computational efficiency and accuracy, ensuring reliable convergence across tested conditions.

\begin{figure}[h!] 
	\centering 
	\includegraphics[width=0.9\textwidth]{Figures/results_cm.png} 
	\caption{Mean number of steps required for convergence as a function of wind speed. Error bars represent the variance of iterations across $1000$ test points.} 
	\label{fig:convergence_steps} 
\end{figure}

\begin{figure}[h!] 
	\centering 
	\includegraphics[width=0.9\textwidth]{Figures/resultsExecutionTime.png} 
	\caption{Average execution time per iteration count over $1000$ tests, measured across $10^5$ sample positions. Error bars indicate variance in execution time.} 
	\label{fig:execution_time} 
\end{figure}

%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%
\afterdoc
%%%%%%%%%%%%%%%
\end{document}
